---
title: "Week 2 - Exercise 2"
author:
  - name: "Gabriela Correia"
    id: jb
    orcid: 0009-0006-8368-7670
    email: g.santoscorreia@students.uu.nl
format: revealjs
logo: cat-face.jpeg
bibliography: bibliography.bib
nocite: "@*"
---

## Cats!

![An animated cat extracted from this [website.](https://www.svgator.com/blog/how-to-create-an-animated-cat/)](cat-logo.png){fig-align="center" width="40%"}


## 
<div style="font-size: 50%;">
```{r}
library(DT)
library(MASS)

data(cats, package="MASS")

data <- data.frame(
  Sex = cats$Sex,
  `Body weight` = cats$Bwt,
  `Heart weight` = cats$Hwt
)

datatable_output <- datatable(data, options = list(pageLength = 5, autoWidth = TRUE), rownames = FALSE)


htmltools::tags$div(
  style = "overflow-y: auto; max-height: 500px;",
  datatable_output
)
```

</div>
This is the cats dataset, from the `MASS` package [@MASS].

---

## Interactive Plot of Cats

```{r}
library(plotly)

fig <- plot_ly(
  data = cats,
  x = ~Bwt,
  y = ~Hwt,
  type = 'scatter',
  mode = 'markers',
  marker = list(size = 8)
) %>%
  layout(
    title = "Cats body and heart weights",
    xaxis = list(title = "Body weight (kg)"),
    yaxis = list(title = "Heart Weight (g)"),
    hovermode = "closest",
    width = 1000,  # Adjust the width as needed
    height = 450, # Adjust the height as needed
    hovermode = "closest"
  )


fig
```

Placing the cursor on top of the datapoints will display the exact X and Y values.


## Two columns + Multi-aligned Equation

:::{.columns}
::: {.column width="50%"}
### Is the cat dead?

![Extracted from [website.](https://www.dreamstime.com/set-physics-formula-symbols-equations-schrodinger-cat-isolated-white-background-hand-drawn-illustration-back-to-school-image274276699)](cat-dead-alive.png){fig-align="center" width="80%"}
:::

::: {.column width="50%"}
### Schrödinger's Equation

$$
\begin{aligned}
  &- \frac{\hbar^2}{2m} \nabla^2 \psi(x, t) + V(x) \psi(x, t) \\
  &= i \hbar \frac{\partial}{\partial t} \psi(x, t)
\end{aligned}
$$
:::
:::

## Code 

Fitting a linear model to see how Heart weight depends on Body weight

```{r}
#| eval: FALSE
#| echo: TRUE



cat_lm <- lm(Hwt ~ Bwt, data = cats)
summary(cat_lm)


```

## Cached and Labeled R Code

````markdown
{r}
#| cache: TRUE
#| label: generated-data

cat_lm <- lm(Hwt ~ Bwt, data = cats)
summary(cat_lm)

````



## Reproducible Environment

Creating a reproducible environment.

```{r}
#| code-line-numbers: true
#| label: renv-envir
#| echo: fenced

library(renv)
renv::init()
```

## References
::: {#refs}
:::